Blocks of extended ancestry around selected loci in hybrid zones

### 2016-02-29 drafts/hybrid_zone.tex
some comments

\abstract{
Hybrid zones formed between recently diverged populations offer an opportunity to better understand the mechanisms underlying reproductive isolation and the process of speciation. \yb{Some form of selection against hybrid genotypes is likely common in many hybrid zones. Here, we use a combination of analytical theory and explicit forward simulations} to gain a better understanding of how this impacts patterns of introgression \yb{across genomic and geographic} space. \alisa{We describe the properties of clines across the genomes as they are formed, and find that locally disadvantageous alleles exist as recent, small families that move relatively rapidly across space}. \yb{We find that, within a population, ancestral blocks surrounding single-locus incompatibilities} are longer than average \yb{ancestry blocks} within the genome, especially when removed from the center of the hybrid zone.  The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.
}

@@

The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
When formerly isolated populations come back into contact 
before reproductive isolation is complete, gene flow is still possible. 
As two such populations interbreed,
individuals' genomes can be labeled according to which ancestral population they are inherited from,
and local interbreeding and migration of such populations
creates a gradient of ancestry assignments across geography,
centered on their point of contact \citep{Barton1985}. 
\plr{(check that neutral clines discussed in barton '85)}\alisa{This is simply stated in Barton1985 without any justification or citation. Can we just cite Sedghifar2015?}
If the populations are sufficiently diverged,
this process leaves a distinct pattern of variation across the genome, 
in which long tracts of divergent haplotypes from each ancestral population  
are broken up by historical recombination events,
leading to substantial correlations in ancestry between distant markers.

Gene flow will eventually erase 
both geographic clines in ancestry, and strong correlations in ancestry.
Subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}. 
The mixing of ancestries of a given genome is mediated by 
the process of recombination, forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003}.   
The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Sedghifar2015}.  
\plr{(check Gravel; isn't that approximate?)}

@@

@@

We expect that, in the absence of selection, blocks of $A$ ancestry across the genome 
will tend to be shorter the further one goes into the $B$ side of the cline,
because they have had more opportunities to recombine with $B$ haplotypes. 
However, we expect that stretches of $A$ ancestry containing a selected site will be longer
than those that do not contain the selected site at the same spatial location,
because lineages containing the selected site have usually been inherited from the $A$ side of the cline recently. 
As discussed above, we expect these lineages move at speed roughly $\sigma/\sqrt{s}$,
so (selected) $A$ alleles at distance $x$ from the cline center
to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago 
(compared to $x^2/\sigma^2$ for a neutral allele),
and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
\plr{(check predicted decay of mean length $1/x$ in simulations?)}
For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.

%%%%%%
\begin{figure}
    \begin{center}
    \end{center}
    \caption{
        \textbf{Haplotype length distributions},
        comparing numerical solutions to simulations.
        At one time, plot 
        probability of seeing a haplotype of length at least $r$ covering that position on the genome,
        against (a) position on the genome, at a few spatial locations;
        or (b) spatial location, at a few positions on the genome.
        \plr{maybe?}
        \label{fig:haplotype_lengths}
    }
\end{figure}

@@

\plr{This needs a bit more precise discussion.}
As expected, we find that $F_B(m,X)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate. 
We find that at the selected locus, $F_B(m,X)$ decreases with X, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}). As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.

@@

\plr{Remove "Future directions", title this "The hybrid filter", and discuss the hybrid filter ideas we had?}
\subsection*{Other patterns of introgression}
\alisa{Could not find a reference to the hybrid filter}
\plr{don't have to call it the hybrid filter, and it looks like maybe I made up that term, but after reading this paper: \citet{martinsen2001hybrid}; these sorts of things are discussed in \citet{Barton1986}}

>>>>

\abstract{
Hybrid zones formed between recently diverged populations offer an opportunity to better understand the mechanisms underlying reproductive isolation and the process of speciation. \yb{Some form of selection against hybrid genotypes is likely common in many hybrid zones. Here, we use a combination of analytical theory and explicit forward simulations} to gain a better understanding of how this impacts patterns of introgression \yb{across genomic and geographic} space. \alisa{We describe the properties of clines across the genomes as they are formed, and find that locally disadvantageous alleles exist as recent, small families that move relatively rapidly across space}. \yb{We find that, within a population, ancestral blocks surrounding single-locus incompatibilities} are longer than average \yb{ancestry blocks} within the genome, especially when removed from the center of the hybrid zone.  The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.
\alisa{ascertainment bias. Tract lengths provide additional source of information, as they provide unbiased information even when used ascertained markers.}

}

@@

The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
When formerly isolated populations come back into contact 
before reproductive isolation is complete, gene flow is still possible. 
As two such populations interbreed,
individuals' genomes can be labeled according to which ancestral population they are inherited from,
and local interbreeding and migration of such populations
creates a gradient of ancestry assignments across geography,
centered on their point of contact \citep{Barton1985}. 
%\plr{(check that neutral clines discussed in barton '85)}
If the populations are sufficiently diverged,
this process leaves a distinct pattern of variation across the genome, 
in which long tracts of divergent haplotypes from each ancestral population  
are broken up by historical recombination events,
leading to substantial correlations in ancestry between distant markers.

Gene flow will eventually erase 
both geographic clines in ancestry, and strong correlations in ancestry.
Subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}. 
The mixing of ancestries of a given genome is mediated by 
the process of recombination, forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003}.   
The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Hellenthal2014,Sedghifar2015}.

@@

\alisa{how small is too small for s? when do we start seeing that peak (1/tau)? Bring paragraph here. Look at smaller spatial scale for the small s simulations}

@@

We expect that, in the absence of selection, blocks of $A$ ancestry across the genome 
will tend to be shorter the further one goes into the $B$ side of the cline,
because they have had more opportunities to recombine with $B$ haplotypes. 
However, we expect that stretches of $A$ ancestry containing a selected site will be longer
than those that do not contain the selected site at the same spatial location,
because lineages containing the selected site have usually been inherited from the $A$ side of the cline recently. 
As discussed above, we expect these lineages move at speed roughly $\sigma/\sqrt{s}$,
so (selected) $A$ alleles at distance $x$ from the cline center
to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago 
(compared to $x^2/\sigma^2$ for a neutral allele),
and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
%\plr{(check predicted decay of mean length $1/x$ in simulations?)}
For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.

%%%%%%
%\begin{figure}
 %   \begin{center}
 %   \end{center}
%    \caption{
  %      \textbf{Haplotype length distributions},
  %      comparing numerical solutions to simulations.
   %     At one time, plot 
   %     probability of seeing a haplotype of length at least $r$ covering that position on the genome,
    %    against (a) position on the genome, at a few spatial locations;
   %     or (b) spatial location, at a few positions on the genome.
   %     \plr{maybe?}
    %    \label{fig:haplotype_lengths}
   % }
%\end{figure}

@@

\plr{This needs a bit more precise discussion.}
\alisa{add motivation-- additional information in haplotype diversity. Present alternate hypotheses. This ties in with our world view within our model}
As expected, we find that the average family size of a $B$ allele decreases with distance into the $A$ side of the hybrid zone. 
% $F_B(m,x)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate. 
We find that at the selected locus, $F_B(m,x)$ decreases with x, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}). As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.

@@

\plr{Remove "Future directions", title this "The hybrid filter", and discuss the hybrid filter ideas we had?}
\subsection*{Adaptive introgression}
\alisa{Could not find a reference to the hybrid filter}
\plr{don't have to call it the hybrid filter, and it looks like maybe I made up that term, but after reading this paper: \citet{martinsen2001hybrid}; these sorts of things are discussed in \citet{Barton1986}}
